Non-kissing and non-crossing complexes \\ for locally gentle algebras

### 2018-07-12 secondPaper/accordionComplexSurfaces.tex

\begin{proof}
The duality provides a bijection~$v\in (Q_\dissection)_0\mapsto v\dual\in (Q_{\dissection\dual})_0$, and also induces a bijection~${\alpha \in (Q_\dissection)_1 \mapsto \alpha\dual \in (Q_{\dissection\dual})_1}$ with~$s(\alpha\dual) = t(\alpha)\dual$ and~$t(\alpha\dual) = s(\alpha)\dual$.
 Moreover, for any two composable arrows~$\alpha,\beta\in(Q_\dissection)_0$, we have~$\alpha\beta\in I_\dissection$ if and only if~$s(\alpha),t(\alpha),t(\beta)$ are three consecutive edges of a~$\dissection$-cell~$f$, if and only if~$s(\alpha)\dual,t(\alpha)\dual,t(\beta)\dual$ are three consecutive edges in~$\dissection\dual$ ending in~$f\dual$, if and only if~$\beta\dual\alpha\dual\notin I_{\dissection\dual}$.
\end{proof}

@@

\begin{remark}
\label{rem:propertiesSurface}
The following observations are useful for the computation of examples.
\begin{enumerate}[(i)]
\item The set~$V_{\bar Q}$ has one vertex for each straight walk in~$\bar Q$ (equivalently, for each maximal path in~$\bar Q$).
      Finite straight walks yield vertices on the boundary of~$\surface_{\bar Q}$, while infinite cyclic straight walks in~$\bar Q$ yield punctures~$\surface_{\bar Q}$ in~$V_{\bar Q}$.
      We denote by~$p$ the number of infinite cyclic straight walks in~$\bar Q$.
\item \label{item:edges}
      The dissection~$\dissection_{\bar Q}$ has one edge for each vertex~$v \in \bar Q_0$, obtained by concatenation of the sides~$\Enrt{\alpha} = \Enrs{\beta}$ and~$\Enrt{\alpha'} = \Enrs{\beta'}$ where~$v = t(\alpha) = s(\beta) = t(\alpha') = s(\beta')$, $\alpha\beta \notin I$ and~$\alpha'\beta' \notin I$. We denote by~$\edgeof(v)$ the edge of~$\dissection_{\bar Q}$ corresponding to~$v$.
\item The dissection~$\dissection_{\bar Q}$ has one $\ell$-cell for each straight walk of length~$\ell$ in~$\bar Q\koszul$.
\item Similar statements hold dually for~${{V\dual}\!\!_{\bar Q}}$ and~${\dissection\dual\!\!_{\bar Q}}$, and the notations~$p\dual$ and~$\dualedgeof(v)$ are defined similarly.
\item The number of punctures of~$\surface_{\bar Q}$ is the number~$p + p\dual$ of infinite straight walks in~$\bar Q$ and in~$\bar Q\koszul$.
\item \label{item:boundary}
      The number~$b$ of boundary components of~$\surface_{\bar Q}$ can be computed as follows.
      There are two natural perfect matchings whose vertices are the blossom vertices of~$\bar Q$: one is obtained by joining the endpoints of each finite straight walk of~$\bar Q$, and the other is obtained similarly from~$\bar Q \koszul$.
      Let~$G$ be the superposition of these two perfect matchings.
      Then the number~$b$ of boundary components of~$\surface_{\bar Q}$ is the number of connected components of~$G$.
\item The genus of the surface~$\surface_{\bar Q}$ is
      \[
       g = \frac{|Q_1| - |Q_0| - b - p - p\dual + 2}{2},
      \]
      where~$b$ is the number of boundary components (see above for a way to compute~$b$) and~$p$ the number of punctures (\ie infinite straight walks in~$\bar Q$ and in~$\bar Q\koszul$).
      Indeed, ${2g = 2-\chi(\surface_{\bar Q})}$, where~$\chi(\surface_{\bar Q})$ is the Euler characteristic of~$\surface_{\bar Q}$.
      The dissection~$\dissection$ provides a cellular decomposition of~$\surface_{\bar Q}$ on which we fill the boundary components with disks.  
      The number of faces of this cellular decomposition is~$b + p\dual$, its number of edges is~$|Q_0|$ and its number of vertices is the number of straight walks in~$\bar Q$, which is equal to~$2|Q_0| - |Q_1| + p$.
      The above formula for~$g$ follows.
\end{enumerate}
\end{remark}

@@

\begin{example}
\label{exm:peakDeepFacets}
We have seen in Definition~\ref{def:straightBending} that each vertex~$v$ of~$Q$ gives rise to a \defn{peak walk}~$v_\peak$ (resp.~\defn{deep walk}~$v_\deep$) with a single peak (resp.~deep) at~$v$ and no other corner.
The set of all such walks forms the \defn{peak facet}~$\set{v_\peak}{v \in Q_0}$ (resp.~the \defn{deep facet}~$\set{v_\deep}{v \in Q_0}$) of the non-kissing complex~$\RNKC$.
As an example of Theorem~\ref{thm:complexesCoincide}, let us now describe the corresponding non-crossing facets of~$\RNCC[\dissection_{\bar Q}, {{\dissection\dual}\!\!_{\bar Q}}]$.

@@

We invite the reader to check that~$\curveof(v_\peak) = \vprevious{\edgeof(v)}$ (resp.~$\curveof(v_\deep) = \vnext{\edgeof(v)}$), where~$\edgeof(v)$ denotes the edge of~$\dissection_{\bar Q}$ corresponding to~$v \in Q_0$ (see Remark~\ref{rem:propertiesSurface}\,\eqref{item:edges}).
Therefore, the peak facet (resp.~the deep facet) can be thought of as the dissection~$\dissection_{\bar Q}$ slightly rotated clockwise (resp.~counterclockwise) on the surface~$\surface_{\bar Q}$.
\end{example}

@@

\begin{remark}
\label{rem:countercurrentOrderOnSurface}
Using Theorem \ref{thm:complexesCoincide}, we can translate the definition of the countercurrent order~$\prec_\alpha$ to the language of~$\dissection$-accordions (resp.~of~$\dissection\dual$-slaloms).
An arrow~$\alpha$ in~$\bar Q\blossom$ corresponds to an angle at a point~$v$ between two edges of~$\dissection$ (resp.~at a point~$v\dual$ between two edges of~$\dissection\dual$).
Two distinct non-kissing walks~$\mu_\star, \nu_\star$ marked at~$\alpha$ in~$\bar Q$ correspond to two distinct non-crossing~$\dissection$-accordions (resp.~$\dissection\dual$-slaloms)~$\gamma$ and~$\delta$ passing in front of this angle.
Then~$\gamma \prec_\alpha \delta$ if~$\gamma$ passes closer to~$v$ (resp.~further to~$v\dual$) than~$\delta$ does.
These interpretations of~$\prec_\alpha$ on the surface where already considered in the case of the disk in~\cite{GarverMcConville, MannevillePilaud-accordion}.
%
\begin{figure}[h]
	\capstart
	\centerline{\includegraphics[scale=1]{orderCurves}}
	\caption{The order~$\mu \prec_\alpha \nu$ on marked walks is the order in which the corresponding curves~$\curveof(\mu)$ and~$\curveof(\nu)$ cross the angles of~$\dissection$ and~$\dissection\dual$ corresponding to~$\alpha$.}
	\label{fig:orderCurves}
\end{figure}
\end{remark}

@@

Let~$\bar Q = (Q,I)$ be a locally gentle bound quiver, and let~$(\b{e}_v)_{v \in Q_0}$ denote the standard basis of~$\R^{Q_0}$.

\begin{definition}
For a multiset~$V \eqdef \{v_1, \dots, v_m\}$ of vertices of~$Q_0$, we denote by~$\multiplicityVector_V \in \R^{Q_0}$ the \defn{multiplicity vector} of~$V$, defined~by
\[
\multiplicityVector_V \eqdef \sum_{i \in [m]} \b{e}_{v_i} = \sum_{v \in Q_0} |\set{i \in [m]}{v_i = v}| \, \b{e}_v.
\]
For a string~$\sigma$ of~$\bar Q$, we define~$\multiplicityVector_\sigma \eqdef \multiplicityVector_{V(\sigma)}$ where~$V(\sigma)$ is the multiset of vertices of~$\sigma$. 
\end{definition}

@@

Our objective is now to interpret the three families of vectors of Section~\ref{subsec:gcdVectorsWalks} in terms of accordions and slaloms on the surface.
According to Remark~\ref{rem:propertiesSurface}\,\eqref{item:edges}, each vertex~$p$ of~$Q$ corresponds to an edge~$\edgeof(p)$ in the dissection~$\dissection_{\bar Q}$ and  to an edge~$\dualedgeof(p) = \edgeof(p)\dual$ in the dual dissection~${\dissection\dual\!\!_{\bar Q}}$.
Recall moreover from Lemma~\ref{lemm:curveOfAWalkIsAccordion} and Proposition~\ref{prop:accordionsSlaloms} that each non-kissing walk~$\omega$ in~$\bar Q$ corresponds to a curve~$\curveof(\omega)$ that is both a $\dissection$-accordion and a $\dissection\dual$-slalom.
The following three statements translate the definitions of $\b{g}$-, $\b{c}$- and $\b{d}$-vectors on the surface.
The proofs are left to the reader.
We note that these interpretations were already used in~\cite{MannevillePilaud-accordion, GarverMcConville} when the surface is a disk.
When~$\dissection$ is a triangulation of~$\surface$, these interpretations coincide with those given in terms of triangulations and laminations for cluster algebras from surfaces by S.~Fomin and D.~Thurston~\cite{FominThurston}.

\begin{proposition}
\label{prop:gVectorsSurface}
For a walk~$\omega$ of~$\bar Q$ and a vertex~$p \in Q_0$, the $p$-th entry of the $\b{g}$-vector~$\gvector{\omega}$ is given~by:
\begin{enumerate}[(i)]
\item The number of times that the $\dissection$-accordion~$\curveof(\omega)$ positively crosses the edge~$\edgeof(p)$ of~$\dissection$ minus the number of times it negatively crosses the edge~$\edgeof(p)$ of~$\dissection$. Here, the crossing is positive (resp.~negative) if the two angles of~$\dissection$ crossed by~$\curveof(\alpha)$ just before and after~$\edgeof(p)$ form a~$\SSS$ (resp.~a~$\ZZZ$).
We note that some crossings are neither positive nor negative.
\item The number of times that the $\dissection\dual$-slalom~$\curveof(\omega)$ positively crosses the edge~$\dualedgeof(p)$ of~$\dissection\dual$ minus the number of times it negatively crosses the edge~$\dualedgeof(p)$ of~$\dissection\dual$. Here, the crossing is positive when~$\curveof(\omega)$ leaves a face~$u\dual$ to enter a face~$v\dual$ with the vertex~$u$ being to its right (and thus with~$v$ being to its left), and the crossing is negative otherwise.
We note that, in this case, each crossing is either positive or negative.
\end{enumerate}
\end{proposition}

\begin{proposition}
\label{prop:cVectorsSurface}
Consider a walk~$\omega$ of~$\bar Q$ in a non-kissing facet~$F \in \RNKC$.
Call \defn{distinguished subcurve} of~$\curveof(\omega)$ in~$\curveof(F)$ the part of~$\curveof(\omega)$ which lies strictly between the two angles of~$\dissection$ (or of~$\dissection\dual$) where~$\curveof(\omega)$ is maximal (see Remark~\ref{rem:countercurrentOrderOnSurface}).
For a vertex~$v \in Q_0$, the $v$-th entry of the $\b{c}$-vector~$\cvector{\omega}{F}$ is up to the sign given by:
\begin{enumerate}[(i)]
\item the number of crossings between the distinguished subcurve of~$\curveof(\omega)$ in~$\curveof(F)$ and the edge~$\edgeof(v)$ of the dissection~$\dissection$,
\item the number of times the distinguished subcurve of~$\curveof(\omega)$ in~$\curveof(F)$ follows the edge~$\dualedgeof(v)$ of the dual dissection~$\dissection\dual$. Here, we say that a curve~$\delta$ follows~$\dualedgeof(v)$ if~$\dualedgeof(v)$ lies on the boundary opposite to~$u$ of a face~$u\dual$ of~$\dissection\dual$ crossed by~$\delta$.
\end{enumerate}
\end{proposition}

@@

\begin{proposition}
\label{prop:dVectorsSurface}
Denote by~$X_\deep \eqdef \set{\curveof(v_\peak)}{v \in Q_0}$ the deep facet of the non-crossing complex obtained by slightly rotating~$\dissection_{\bar Q}$ counterclockwise as described in Example~\ref{exm:peakDeepFacets}.
For a walk~$\omega$ of~$\bar Q$ and a vertex~$v \in Q_0$, the $v$-th entry of the $\b{d}$-vector~$\dvector{\omega}$ is either~$-1$ if~$\omega = v_\peak$ or the number of crossings of the curve~$\gamma(\omega)$ with the edge~$\curveof(v_\peak)$ of~$X_\deep$ corresponding to~$v$.
\end{proposition}

@@

\begin{proposition}
\label{rem:signCoherence}
The $\b{g}$-, $\b{c}$-, and~$\b{d}$-vectors have the \defn{sign-coherence property}: for any non-kissing facet~$F \in \RNKC$,
\begin{itemize}
\item for~$v \in Q_0$, the~$v$-th coordinates of all $\b{g}$-vectors~$\gvector{\omega}$ for~$\omega \in F$ have the same sign,
\item for any~$\omega \in F$, all coordinates of the $\b{c}$-vector~$\cvector{\omega}{F}$ have the same sign,
\item for any~$\omega \in F$, all coordinates of the $\b{d}$-vector~$\dvector{\omega}$ have the same sign.
\end{itemize}
%In particular, a $\b{c}$-vector is called \defn{positive} when all its coordinates are non-negative and \defn{negative} when all its coordinates are non-positive.
\end{proposition}

\begin{proof}
For the $\b{g}$-vectors, observe that two walks~$\omega$ and~$\omega'$ whose $\b{g}$-vectors~$\gvector{\omega}$ and~$\gvector{\omega'}$ have opposite signs on their~$v$-th coordinate automatically kiss at~$v$.
For $\b{c}$- and $\b{d}$-vectors, the statement is immediate from the definition.
\end{proof}

>>>>

\begin{proof}
The duality provides a bijection~$a\in (Q_\dissection)_0\mapsto a\dual\in (Q_{\dissection\dual})_0$, and also induces a bijection~${\alpha \in (Q_\dissection)_1 \mapsto \alpha\dual \in (Q_{\dissection\dual})_1}$ with~$s(\alpha\dual) = t(\alpha)\dual$ and~$t(\alpha\dual) = s(\alpha)\dual$.
 Moreover, for any two composable arrows~$\alpha,\beta\in(Q_\dissection)_0$, we have~$\alpha\beta\in I_\dissection$ if and only if~$s(\alpha),t(\alpha),t(\beta)$ are three consecutive edges of a~$\dissection$-cell~$f$, if and only if~$s(\alpha)\dual,t(\alpha)\dual,t(\beta)\dual$ are three consecutive arcs in~$\dissection\dual$ ending in~$f\dual$, if and only if~$\beta\dual\alpha\dual\notin I_{\dissection\dual}$.
\end{proof}

@@

\begin{remark}
\label{rem:propertiesSurface}
The following observations are useful for the computation of examples.
\begin{enumerate}[(i)]
\item The set~$V_{\bar Q}$ has one vertex for each straight walk in~$\bar Q$ (equivalently, for each maximal path in~$\bar Q$).
      Finite straight walks yield vertices on the boundary of~$\surface_{\bar Q}$, while infinite cyclic straight walks in~$\bar Q$ yield punctures~$\surface_{\bar Q}$ in~$V_{\bar Q}$.
      We denote by~$p$ the number of infinite cyclic straight walks in~$\bar Q$.
\item \label{item:edges}
      The dissection~$\dissection_{\bar Q}$ has one edge for each vertex~$a \in \bar Q_0$, obtained by concatenation of the sides~$\Enrt{\alpha} = \Enrs{\beta}$ and~$\Enrt{\alpha'} = \Enrs{\beta'}$ where~$a = t(\alpha) = s(\beta) = t(\alpha') = s(\beta')$, $\alpha\beta \notin I$ and~$\alpha'\beta' \notin I$. We denote by~$\edgeof(a)$ the edge of~$\dissection_{\bar Q}$ corresponding to~$a$.
\item The dissection~$\dissection_{\bar Q}$ has one $\ell$-cell for each straight walk of length~$\ell$ in~$\bar Q\koszul$.
\item Similar statements hold dually for~${{V\dual}\!\!_{\bar Q}}$ and~${\dissection\dual\!\!_{\bar Q}}$, and the notations~$p\dual$ and~$\dualedgeof(a)$ are defined similarly.
\item The number of punctures of~$\surface_{\bar Q}$ is the number~$p + p\dual$ of infinite straight walks in~$\bar Q$ and in~$\bar Q\koszul$.
\item \label{item:boundary}
      The number~$b$ of boundary components of~$\surface_{\bar Q}$ can be computed as follows.
      There are two natural perfect matchings whose vertices are the blossom vertices of~$\bar Q$: one is obtained by joining the endpoints of each finite straight walk of~$\bar Q$, and the other is obtained similarly from~$\bar Q \koszul$.
      Let~$G$ be the superposition of these two perfect matchings.
      Then the number~$b$ of boundary components of~$\surface_{\bar Q}$ is the number of connected components of~$G$.
\item The genus of the surface~$\surface_{\bar Q}$ is
      \[
       g = \frac{|Q_1| - |Q_0| - b - p - p\dual + 2}{2},
      \]
      where~$b$ is the number of boundary components (see above for a way to compute~$b$) and~$p$ the number of punctures (\ie infinite straight walks in~$\bar Q$ and in~$\bar Q\koszul$).
      Indeed, ${2g = 2-\chi(\surface_{\bar Q})}$, where~$\chi(\surface_{\bar Q})$ is the Euler characteristic of~$\surface_{\bar Q}$.
      The dissection~$\dissection$ provides a cellular decomposition of~$\surface_{\bar Q}$ on which we fill the boundary components with disks.  
      The number of faces of this cellular decomposition is~$b + p\dual$, its number of edges is~$|Q_0|$ and its number of vertices is the number of straight walks in~$\bar Q$, which is equal to~$2|Q_0| - |Q_1| + p$.
      The above formula for~$g$ follows.
\end{enumerate}
\end{remark}

@@

\begin{example}
\label{exm:peakDeepFacets}
We have seen in Definition~\ref{def:straightBending} that each vertex~$a$ of~$Q$ gives rise to a \defn{peak walk}~$a_\peak$ (resp.~\defn{deep walk}~$a_\deep$) with a single peak (resp.~deep) at~$a$ and no other corner.
The set of all such walks forms the \defn{peak facet}~$\set{a_\peak}{a \in Q_0}$ (resp.~the \defn{deep facet}~$\set{a_\deep}{a \in Q_0}$) of the non-kissing complex~$\RNKC$.
As an example of Theorem~\ref{thm:complexesCoincide}, let us now describe the corresponding non-crossing facets of~$\RNCC[\dissection_{\bar Q}, {{\dissection\dual}\!\!_{\bar Q}}]$.

@@

We invite the reader to check that~$\curveof(a_\peak) = \vprevious{\edgeof(a)}$ (resp.~$\curveof(a_\deep) = \vnext{\edgeof(a)}$), where~$\edgeof(a)$ denotes the edge of~$\dissection_{\bar Q}$ corresponding to~$a \in Q_0$ (see Remark~\ref{rem:propertiesSurface}\,\eqref{item:edges}).
Therefore, the peak facet (resp.~the deep facet) can be thought of as the dissection~$\dissection_{\bar Q}$ slightly rotated clockwise (resp.~counterclockwise) on the surface~$\surface_{\bar Q}$.
\end{example}

@@

\begin{remark}
\label{rem:countercurrentOrderOnSurface}
Using Theorem \ref{thm:complexesCoincide}, we can translate the definition of the countercurrent order~$\prec_\alpha$ to the language of~$\dissection$-accordions (resp.~of~$\dissection\dual$-slaloms).
An arrow~$\alpha$ in~$\bar Q\blossom$ corresponds to an angle at a point~$v$ between two arcs of~$\dissection$ (resp.~at a point~$f\dual$,~$v\in f$, between two edges of~$\dissection\dual$).
Two distinct non-kissing walks~$\mu_\star, \nu_\star$ marked at~$\alpha$ in~$\bar Q$ correspond to two distinct non-crossing~$\dissection$-accordions (resp.~$\dissection\dual$-slaloms)~$\gamma$ and~$\delta$ passing in front of this angle.
Then~$\gamma \prec_\alpha \delta$ if~$\gamma$ passes closer to~$v$ (resp.~further to~$f\dual$) than~$\delta$ does.
These interpretations of~$\prec_\alpha$ on the surface where already considered in the case of the disk in~\cite{GarverMcConville, MannevillePilaud-accordion}.
%
\begin{figure}[h]
	\capstart
	\centerline{\includegraphics[scale=1]{orderCurves}}
	\caption{The order~$\mu \prec_\alpha \nu$ on marked walks is the order in which the corresponding curves~$\curveof(\mu)$ and~$\curveof(\nu)$ cross the angles of~$\dissection$ and~$\dissection\dual$ corresponding to~$\alpha$.}
	\label{fig:orderCurves}
\end{figure}
\end{remark}

@@

Let~$\bar Q = (Q,I)$ be a locally gentle bound quiver, and let~$(\b{e}_a)_{a \in Q_0}$ denote the standard basis of~$\R^{Q_0}$.

\begin{definition}
Let~$V \eqdef \{a_1, \dots, a_m\}$ be a multiset of vertices of~$Q_0$. We denote by~$\multiplicityVector_V \in \R^{Q_0}$ the \defn{multiplicity vector} of~$V$, defined~by:
\[
\multiplicityVector_V \eqdef \sum_{i \in [m]} \b{e}_{a_i} = \sum_{a \in Q_0} |\set{i \in [m]}{a_i = a}| \, \b{e}_a.
\]
For a string~$\sigma$ of~$\bar Q$, we define~$\multiplicityVector_\sigma \eqdef \multiplicityVector_{V(\sigma)}$ where~$V(\sigma)$ is the multiset of vertices of~$\sigma$. 
\end{definition}

@@

Our objective is now to interpret the three families of vectors of Section~\ref{subsec:gcdVectorsWalks} in terms of accordions and slaloms on the surface.
According to Remark~\ref{rem:propertiesSurface}\,(\eqref{item:edges}), each vertex~$a$ of~$Q$ corresponds to an edge~$\edgeof(a)$ in the dissection~$\dissection_{\bar Q}$ and  to an edge~$\dualedgeof(a) = \edgeof(a)\dual$ in the dual dissection~${\dissection\dual\!\!_{\bar Q}}$.
Recall moreover from Lemma~\ref{lemm:curveOfAWalkIsAccordion} and Proposition~\ref{prop:accordionsSlaloms} that each non-kissing walk~$\omega$ in~$\bar Q$ corresponds to a curve~$\curveof(\omega)$ that is both a $\dissection$-accordion and a $\dissection\dual$-slalom.
The following three statements translate the definitions of $\b{g}$-, $\b{c}$- and $\b{d}$-vectors on the surface.
The proofs are left to the reader.
We note that these interpretations were already used in~\cite{MannevillePilaud-accordion, GarverMcConville} when the surface is a disk.
When~$\dissection$ is a triangulation of~$\surface$, these interpretations coincide with those given in terms of triangulations and laminations for cluster algebras from surfaces by S.~Fomin and D.~Thurston~\cite{FominThurston}.

\begin{proposition}
\label{prop:gVectorsSurface}
For a walk~$\omega$ of~$\bar Q$ and a vertex~$a \in Q_0$, the $a$-th entry of the $\b{g}$-vector~$\gvector{\omega}$ is given~by:
\begin{enumerate}[(i)]
\item The number of times that the $\dissection$-accordion~$\curveof(\omega)$ positively crosses the edge~$\edgeof(a)$ of~$\dissection$ minus the number of times it negatively crosses the edge~$\edgeof(a)$ of~$\dissection$. Here, the crossing is positive (resp.~negative) if the two angles of~$\dissection$ crossed by~$\curveof(\alpha)$ just before and after~$\edgeof(a)$ form a~$\SSS$ (resp.~as a~$\ZZZ$).
We note that some crossings are neither positive nor negative.
\item The number of times that the $\dissection\dual$-slalom~$\curveof(\omega)$ positively crosses the edge~$\dualedgeof(a)$ of~$\dissection\dual$ minus the number of times it negatively crosses the edge~$\dualedgeof(a)$ of~$\dissection\dual$. Here, the crossing is positive when~$\curveof(\omega)$ leaves a face~$u\dual$ to enter a face~$v\dual$ with the vertex~$u$ being to its right (and thus with~$v$ being to its left), and the crossing is negative otherwise.
We note that, in this case, each crossing is either positive or negative.
\end{enumerate}
\end{proposition}

\begin{proposition}
\label{prop:cVectorsSurface}
Consider a walk~$\omega$ of~$\bar Q$ in a non-kissing facet~$F \in \RNKC$.
Call \defn{distinguished subcurve} of~$\curveof(\omega)$ in~$\curveof(F)$ the part of~$\curveof(\omega)$ which lies strictly between the two angles of~$\dissection$ (or of~$\dissection\dual$) where~$\curveof(\omega)$ is maximal (see Remark~\ref{rem:countercurrentOrderOnSurface}).
For a vertex~$a \in Q_0$, the $a$-th entry of the $\b{c}$-vector~$\cvector{\omega}{F}$ is up to the sign given by:
\begin{enumerate}[(i)]
\item the number of crossings between the distinguished subcurve of~$\curveof(\omega)$ in~$\curveof(F)$ and the edge~$\edgeof(a)$ of the dissection~$\dissection$,
\item the number of times the distinguished subcurve of~$\curveof(\omega)$ in~$\curveof(F)$ follows the edge~$\dualedgeof(a)$ of the dual dissection~$\dissection\dual$. Here, we say that a curve~$\delta$ follows~$\dualedgeof(a)$ if~$\dualedgeof(a)$ lies on the boundary opposite to~$u$ of a face~$u\dual$ of~$\dissection\dual$ crossed by~$\delta$.
\end{enumerate}
\end{proposition}

@@

\begin{proposition}
\label{prop:dVectorsSurface}
Denote by~$X_\deep \eqdef \set{\curveof(a_\peak)}{a \in Q_0}$ the deep facet of the non-crossing complex obtained by slightly rotating~$\dissection_{\bar Q}$ counterclockwise as described in Example~\ref{exm:peakDeepFacets}.
For a walk~$\omega$ of~$\bar Q$ and a vertex~$a \in Q_0$, the $a$-th entry of the $\b{d}$-vector~$\dvector{\omega}$ is either~$-1$ if~$\omega = a_\peak$ or the number of crossings of the curve~$\gamma(\omega)$ with the edge~$\curveof(a_\peak)$ of~$X_\deep$ corresponding to~$a$.
\end{proposition}

@@

\begin{proposition}
\label{rem:signCoherence}
The $\b{g}$-, $\b{c}$-, and~$\b{d}$-vectors have the \defn{sign-coherence property}: for any non-kissing facet~$F \in \RNKC$,
\begin{itemize}
\item for~$a \in Q_0$, the~$a$-th coordinates of all $\b{g}$-vectors~$\gvector{\omega}$ for~$\omega \in F$ have the same sign,
\item for any~$\omega \in F$, all coordinates of the $\b{c}$-vector~$\cvector{\omega}{F}$ have the same sign,
\item for any~$\omega \in F$, all coordinates of the $\b{d}$-vector~$\dvector{\omega}$ have the same sign.
\end{itemize}
%In particular, a $\b{c}$-vector is called \defn{positive} when all its coordinates are non-negative and \defn{negative} when all its coordinates are non-positive.
\end{proposition}

\begin{proof}
For the $\b{g}$-vectors, observe that two walks~$\omega$ and~$\omega'$ whose $\b{g}$-vectors~$\gvector{\omega}$ and~$\gvector{\omega}$ have opposite signs on their~$a$-th coordinate automatically kiss at~$a$.
For $\b{c}$- and $\b{d}$-vectors, the statement is immediate from the definition.
\end{proof}
